---
title: "Snowshoe Report - Emily Pelstring"
author: "Emily Pelstring"
date: "11/6/2021"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(ggthemes)
library(lubridate)
library(ggbeeswarm)
library(kableExtra)
library(effsize)
library(broom)
```

### 1. Introduction

This report provides an exploratory overview of juvenile snowshoe hare (Lepus americanus) populations in the Bonanza Creek Long Term Ecological Research SIte (Bonanza Creek LTER, 2021), Alaska at three different sampling locations from 1999 to 2012 (Kielland et al. 2017). Several components were looked at regarding the hare population, the first being juvenile hare trap count by year. Then, the weight of juvenile hares were compared with regards to sex and collection site. Finally, the difference between male and female juvenile hare weights are compared using descriptive statistics and means comparisons. 

### 2. Data and Methods

Physical data about juvenile snowshoe hares was made available by Knut Kielland and colleagues at 5 locales in the Tanana valley, Alaska. The data contains observations for 3197 hares, juvenile and adult, collected from the years of 1999 to 2012. Following exploratory visualization with graphs including a column graph, beeswarm graph, and scatter-plot, mean weights of juvenile male and female hares were compared by two-sample t-tests using a significance level ($\alpha$) of 0.05 throughout. Differences between groups are described by Cohen's *d* effect size. The relationship between weight and hind foot length is explored by simple linear regression and Pearson's r correlation. All analyses are in R Studio version 1.4.1717 and R version 4.1.1.

### 3. Preliminary Results

#### A. Juvenile Hare Initial Investigation

```{r fig.align = 'center'}
## Reading in the Data 
bonanza_hares <- read.csv(here("data", "bonanza_hares.csv"))

## Creating a column to include years 
bonanza_hares <- bonanza_hares %>%
  mutate(date_new = mdy(date)) %>%
  mutate(year_new = year(date_new))

## total hare trap counts 
total_hare_trapcounts <- bonanza_hares %>%
  group_by(year_new) %>%
  filter(age == "j") %>%
  count()

## Mean and Median of Count
mean_count <- round(mean(total_hare_trapcounts$n), 1)
median_count <- round(median(total_hare_trapcounts$n), 1)

## visualization of total hare trap counts 
total_hare_trapcounts %>%
  ggplot(aes(x = year_new,
             y = n)) +
  geom_col() + 
  theme_economist() +
  theme(text = element_text(size = 11, family = "Times New Roman")) +
  labs(x = "Year",
       y = "Number of Hares",
       title = "Total Juvenile Hare Trap Counts by Year (PELSTRING)")

```

**Figure 1.** Total number of juvenile hares observed from the years 1999 - 2012. Each bar signifies total number of juvenile hares (y-axis) observed by year (x-axis). Data: (Kielland et al. 2017).

##### Major Findings 

Among reporting years, the highest number of juvenile hare observations was in 1999 with 126 observations, while the lowest took place in 2010 with 2 observations. The mean number of juvenille hare observations per year is `r mean_count` hares, and the median number of juvenile hare observations through all years is `r median_count` hares. According to the column graph, it appears as though there was a marked decline in juvenile hare populations from when observations first began in 1999. This decline in juvenile hare observations may signal that there is a decline in reproduction rates of adult hares; however, before these assumptions can be made, sampling methodology must be examined. Dividing each observation by number of days spent recording, as well as number of traps set out, would help to normalize the data between years. 

#### B. Visualization of Juvenile Hare Weights by Site and Sex

```{r fig.align = 'center'}
## Filtering for only juvenile hares 
juvenile_hares <- bonanza_hares %>%
  filter(age == "j") %>% 
  mutate(grid = case_when(
    grid %in% c("bonrip") ~ "Riparian",
    grid %in% c("bonbs") ~ "Black Spruce",
    grid %in% c("bonmat") ~ "Mature Mixed"))

## Plotting for Juvenile hare weights by site and sex. 
ggplot(data = juvenile_hares, aes(x = sex, y = weight)) + #selecting data, setting x value as site and y value as weight 
  geom_beeswarm(aes(color = sex)) + #type of graph, splitting color by sex
  scale_color_manual(values = c("coral","blue3","coral"), #adding specific colors by vector 
                     name = "Sex", labels = c("female", "male", "NA")) + #changing label text
  theme_economist() + #adding theme
  theme(text = element_text(size = 10, family = "Times New Roman")) + #changing graph text size and font
  labs(x = "", #labeling axes
       y = "Weight (grams)",#labeling axes
       title = "Weight of Juvenile Hares by Sex and Site (PELSTRING)") + #adding title
  facet_wrap(~grid) #faceting by site 
```
**Figure 2.** Distribution of weight of juvenile hares at three sampling sites, separated by sex via coloring. Sex is displayed on the x-axis while weight is displayed on the y-axis, and the data is faceted by site. The amount of observations between sites varied, with Bonanza Back Spruce having the lowest number of observations available and Bonanza Riparian with the highest number of samples available. Data: (Kielland et al. 2017).

##### Main Takeaways 

According to the graph, it appears that there is some difference between male and female juvenile hare weights at the three observation sites. It appears as though the weight of male juvenile hares is slightly higher than the weight of female juvenile hares. There is no significant difference in distribution of weights between the three sites. There are fewer observations available at the Black Spruce site compared to the Mature Mixed and Riparian sites, which may impact the ability to draw conclusions regarding the lack of significant difference in weights between that and the two other sites. 

#### C. Juvenile Weight Comparison (male & female snowshoe hares)

**Table 1.** Descriptive statistics (mean, standard deviation, sample size) for male and female juvenile hares. Data: (Kielland et al. 2017).
```{r fig.align = 'center'}
juvenile_table <- juvenile_hares %>% # creating new data set
  group_by(sex) %>% #grouping by sex
  mutate(sex = case_when( #mutating to change the labels of f/m to female/male
    sex %in% c("f") ~ "Female",
    sex %in% c("m") ~ "Male",
    sex %in% c("NA") ~ "NA")) %>%
  summarize(
    mean_juvenile_weight = round(mean(weight, na.rm = TRUE), 1), #finding the mean of juvenile weight, rounding to 1 decimal place
    sd_juvenile_weight= round(sd(weight, na.rm = TRUE), 1), #finding the standard deviation of juvenile weight, rounding to 1 decimal 
    n = n()) #adding sample size

juvenile_table %>% # creating a table
  kable(col.names = c("Sex", "Mean Juvenile Weight (g)", "Standard Deviation", "Sample Size")) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
   
   
```
```{r include = FALSE}
## Exploratory Graphs 

ggplot(data = juvenile_hares, #selecting the data
       aes(x = weight)) + #setting x axis 
  geom_histogram(bins = 10) + #choosing histogram plot
  facet_wrap(~sex) #faceting by sex
  
ggplot(data = juvenile_hares, #selecting the data
       aes(sample = weight)) + #setting x axis
  geom_qq() + #choosing qq plot
  facet_wrap(~sex) #faceting by sex
```



##### Means Comparisons 

```{r}
# Finding difference in means 

difference_in_means <- juvenile_table$mean_juvenile_weight[2] - juvenile_table$mean_juvenile_weight[1]

# storing vectors for male and female weights for use in Cohen and T-test

female_weight <- juvenile_hares %>% #vector for female weights
  filter(sex == "f") %>% #filtering for only female
  drop_na(weight) %>% #removing na values
  pull(weight) #vector of just the weight column

male_weight <- juvenile_hares %>% #vector for male weights
  filter(sex == "m") %>% #filtering for only male
  drop_na(weight) %>% #removing na values
  pull(weight) #vector of just the weight column

weight_t <- t.test(x = male_weight, y = female_weight) # t-test

# weight_t #viewing t-test

weight_cohen <- cohen.d(female_weight, male_weight) #effect size 

# weight_cohen #viewing cohen 

```

Mean weight for male juvenile hares (`r round(weight_t$estimate[1],2)` g) is greater than that for female juvenile hares (`r round(weight_t$estimate[2],2)` g); The difference in the means between them is `r difference_in_means` g. The effect size is small (Cohen's *d* = `r round(weight_cohen$estimate,2)`) and the difference significant (two-sample t-test, t = `r round(weight_t$statistic,2)`, p = 0.007)

#### D. Juvenile Hare Weight versus Hind Foot Length 

```{r fig.align = 'center'}
ggplot(data = juvenile_hares, #selecting for the data
       aes(x = weight, y = hindft)) + #setting x and y axis
  geom_jitter() + #creating scatterplot 
  geom_smooth(method = "lm") + #adding lm 
  theme_economist() + #choosing theme
  theme(text = element_text(size = 11, family = "Times New Roman")) + #changing font and size
  labs(x = "Weight (g)", #adding labels
       y = "Hind Foot Length (mm)",
       title = "Weight Versus Hind Foot Length (PELSTRING)") #title
```
**Figure 3.** Distribution of weight (in grams) of juvenile hares versus hind foot length (in millimeters). Juvenile hare weight is displayed on the x-axis, while hind foot length is on the y-axis. Each hare observation is represented by a dot on the graph. Data: (Kielland et al. 2017).

```{r}
## Finding Pearson Coefficient and LM

hare_cor <- cor.test(juvenile_hares$weight, juvenile_hares$hindft) # pearson coefficient 

hare_cor_tidy <- tidy(hare_cor) #tidy pearson

# hare_cor_tidy #viewing pearson

hare_lm <- lm(weight ~ hindft, data = juvenile_hares) #finding lm

hare_lm_glance <- glance(hare_lm) #glancing at lm

# hare_lm_glance #viewing lm

hare_lm_tidy <- tidy(hare_lm) #tidying lm

# hare_lm_tidy #viewing tidy lm

```
The relationship between juvenile hare weight and hind foot length appears relatively linear (Figure 3). Linear regression revealed that weight significantly predicts hind foot length (p < 0.001, R^2^ = `r round(hare_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(hare_lm_tidy$estimate[2], 2)` g mm^-1^. Weight and hind foot length are moderately, and significantly, positively correlated (Pearson's *r* = `r round(hare_cor_tidy$estimate,2)`, p < 0.001). One issue with linear regression in this instance is that homoscedasticity is not satisfied. I do not think this linear regression will represent the relationship between hind foot length and weight in the best way, given that the variance in residuals does not seem homogeneous throughout the spread of samples. 

### 4. Summary

This data analysis reveals these initial findings: 

- The total juvenile hare trap counts by year steadily declined from the beginning of sampling (1999) to the end (2012).
- There is a statistically significant difference in means between male and female juvenile snowshoe hare weights and the effect size is small. 
- The relationship between juvenile hare weight and hind foot length is relatively linear (by simple linear regression), with a slope (expected average increase in hind foot length (in mm) with a 1 gram increase of body weight) of `r round(hare_lm_tidy$estimate[2], 2)` g mm^-1^. 

### 5. Citation 

**Data citation:** Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14
