---
title: "Snowshoe Hares - Bonanza Creek Experimental Project"
author: "Shayan Kaveh"
date: "11/6/2021"
output: html_document
---

## Overview

**Data citation:** Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. <https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14>

## Set Up

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center', out.extra = 'angle=90')
library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(kableExtra)
library(plotly)
library(ggbeeswarm)
```

for analysis: assume normality & homosedaciticity 

## Read in data

```{r}
hares <- read_csv(here("data", "bonanza_hares.csv")) %>% 
  # Clean up data to create dates column
  mutate(mdy = mdy(date)) %>% 
  mutate(year = year(mdy)) %>% 
  filter(age == "j") %>% 
  mutate(site = case_when(
    grid == "bonrip" ~ "Bonrip",
    grid == "bonmat" ~ "Bonmat",
    grid == "bonbs" ~ "Bonbs"))

# Make a juvenile hare data set

j_hares <- hares %>% 
  filter(age == "j") %>% 
  mutate(site = case_when(
    grid == "bonrip" ~ "Bonrip",
    grid == "bonmat" ~ "Bonmat",
    grid == "bonbs" ~ "Bonbs")) %>% 
  group_by(year) %>% 
  summarize(count = n())
```

## Annual Juvenile Hair Trap Counts
```{r}
ggplot(data = hares, aes(x = year)) +
  geom_histogram(bins = 15, color = "#8e715f", fill = "#fcefe4") +
  # this is so extra, but my outline color is that of a snowshoe hare when it's brown and the fill color is their fur color in the winter... lol
  scale_x_continuous(breaks = 1999:2012) +
  theme_classic() +
  labs(title = "Juvenile Snowshoe Hare Trap Count By Year",
       x = "Year",
       y = "Number of Hares") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```

## Summarize these statistics

```{r}
j_hares_summary <- j_hares %>% 
  summarize(mean_count = mean(count, na.rm = TRUE),
            median_count = median(count, na.rm = TRUE),
            sd_count = round(sd(count, na.rm = TRUE),1),
            max_count = max(count),
            min_count = min(count))

j_hares_summary %>% 
  kable(col.names = c("Mean Count",
                      "Median Count",
                      "Standard Deviation",
                      "Maximum Count",
                      "Minimum Count"),
        caption = "<center>Table 1. Juvenile Hare Trap Count Statistics<center>",
        digits = 2) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
```

Takeaways:

## Juvenile Hare Weights
Create a finalized data visualization (or visualizations) in which you compare **juvenile hare weights by sex and site (grid)** (ignore other variables). You should include groups where these variables are not recorded (`NA`). Include finalized figure captions below your visualization(s).

Then, in 2 - 3 sentences below your visualization(s) describe in text the main takeaways that you would want the audience to understand from your figures.

Note: There are multiple ways to visualize these comparisons. We encourage you to think carefully then make decisions about how to visualize these comparisons, including:

Can I use faceting to simplify the comparisons into manageable pieces?
Am I showing the audience the data, summary statistics, or both?
Have I updated graph labels, headers in facets, etc. to make it as straightforward as possible for viewers?
Have I simplified and finalized my data visualization(s) to include components that are useful for the audience?

```{r}
gender_j_hares <- hares %>% 
  mutate(sex = case_when(
    sex == "m" ~ "Male",
    sex == "f" ~ "Female"))

ggplot(data = gender_j_hares, 
       aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex)) +
  scale_color_manual(values = c("darkseagreen","lightcoral","indianred1")) +
  geom_boxplot(fill = NA, width = 0.2, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  theme_grey() +
  facet_wrap(~site) +
  labs(x = "Sex",
       y = "Weight (g)",
       title = "Distribution of Juvenile Snowshoe Hare Weights by Sex",
       color = "Sex") + # "color = " changes legend title 
  theme(plot.title = element_text(hjust = 0.5),
        legend.key = element_rect(fill = "white"))
```

